EDIpack2: interoperable Lanczos-based solver for generic quantum impurity problems

### 2025-05-07 03_edipack2.tex
GB: heavy rewrite of the iRDM (3.10) section
Summation indices were misleading, the normal case was written (in the formula) to be diagonal, which is not for Norb > 1. The listing were correct, so it was just a matter of notational mess. I also added a formal proof to why the normal iRDM is block-diagonal with respect to impurity Nup and Ndw sectors.
It should read pretty smooth now.

I also edited many sub- and super-scripts, adding the usual \mathrm{}, when they are words (bath, imp, int, e-ph, ...)

Finally, I did some light editing on the Attractive Hubbard example, just because I found so

\subsection{The quantum impurity problem}\label{sSecQIM}
We consider a general quantum impurity problem defined by the
following Hamiltonian:
$$
\hat{H} = \hat{H}_{imp} + \hat{H}_{bath} + \hat{H}_{hyb} + \hat{H}_\mathrm{ph} + \hat{H}_{e-ph}
$$
which describes a  multi-orbital interacting quantum
impurity coupled to an electronic bath and to local, i.e. Holstein,
phonons. 
\begin{equation}\label{Himp}
  \begin{split}
    \hat{H}_{imp} & = \hat{H}^0_{imp} + \hat{H}^{int}_{imp}\\
    \hat{H}^0_{imp} & =
    \sum_{\a\b\sigma\sigma'}h^{0}_{\a\b\sigma\sigma'}d^{+}_{\a\sigma}d_{\b\sigma'}\\
\end{split}
\end{equation}

@@

\begin{equation}\label{Hint}
  \begin{split}
    \hat{H}^{int}_{imp} &=U\sum_{\a}n_{\a\uparrow}n_{\a\downarrow}+U'\sum_{\a\neq \b}n_{\a\uparrow}n_{\b\downarrow}+(U'-J)\sum_{\a<\b,\sigma}n_{\a\sigma}n_{\b\sigma}\\
    &{\phantom =}- J_X\sum_{\a\neq
      \b}d^{+}_{\a\uparrow}d_{\a\downarrow}d^{+}_{\b\downarrow}d_{\b\uparrow}+J_P\sum_{\a
      \neq
      \b}d^{+}_{\a\uparrow}d^{+}_{\a\downarrow}d_{\b\downarrow}d_{\b\uparrow}\\
\end{split}
\end{equation}

where $d_{\a\sigma}$ ($d^+_{\a\sigma}$) are the destruction (creation)
second-quantization operators for impurity electrons in the
orbital $\alpha=1,\dots,N_\a$, with $N_\a$ the number of orbitals,
with spin $\sigma=\up,\dw$ and whose occupation is described
by the  operator  
$n_{\alpha\sigma}=d^{+}_{\alpha\sigma}d_{\alpha\sigma}$. 
The non-interacting internal structure of the impurity is described by
the $h^{0}_{\a\b\sigma\sigma'}$ matrix. 
$\hat{H}^{int}$ describes the  local multi-orbital
interaction~\cite{Georges2013ACMP} which, for simplicity, we take as a
generalized Hubbard-Kanamori form.  
The first three terms represent the density-density part of the
interaction, where $U$ is the local intra-orbital Coulomb repulsion,
$U'$ the inter-orbital one and $J$ the Hund's coupling~\cite{Georges2013ACMP}.  
The  remaining two terms are, respectively, the spin-exchange and
pair-hopping which we considered with their respective independent
couplings $J_X$ and $J_P$.
In the case $N_\a=3$ a fully symmetric $SU(3)_{orbital}\times SU(2)_{spin}\times
U(1)_{charge}$ form of the interaction is obtained by setting $U'=U-2J$ and
$J_X=J_P=J$~\cite{Georges2013ACMP}. Different choices, preserving part
of the combined symmetry group, can be made for other values of
$N_\a$~\cite{Georges2013ACMP}.

The bath part and its coupling to the impurity has the form: 
\begin{equation}\label{Hbath}
  \begin{split}
    \hat{H}_{bath} &=
    \sum_p\sum_{\a\b\sigma\sigma'}h^p_{\a\b\sigma\sigma'}a^{+}_{p\a\sigma}a_{p\b\sigma'}\\
    %
    \hat{H}_{hyb} &= \sum_p\sum_{\a\b\sigma\sigma'}V^p_{\a\b\sigma\sigma'}d^{+}_{\a\sigma}a_{p\b\sigma'}+H.c. \\
\end{split}
\end{equation}
where $p=1,\dots,N_\mathrm{bath}$ is an index running over a finite number of
bath elements, $a_{p\alpha\sigma}$ ($a^+_{p\alpha\sigma}$) are the destruction (creation) operators for
the bath electrons with index $p$, with orbital $\alpha$ and
spin $\sigma$.
The  properties of each bath level are described by the
matrices $h^p_{\a\b\sigma\sigma'}$. As such any bath element can be
composed of several electronic levels according the bath topology,
which will be discussed further in the following.
Each bath level couples to the impurity with an amplitude
$V^p_{\a\b\s\sigma'}$ which we allow to couple different orbital and
opposite spins.

Finally, the electron-phonon part of the quantum impurity problems is
described by the Hamiltonian terms: 
\begin{equation}\label{Hph}
  \begin{split}
    \hat{H}_\mathrm{ph}&=\sum_q \omega_{0q} b_q^+b_q\\
    % 
    \hat{H}_{e-ph} &= \sum_q\sum_{\a\b\sigma} g_{\a\b} d^+_{\a\sigma}d_{\b\sigma}(b_q+b_q^+)
\end{split}
\end{equation}
where $q=1,\dots,N_q$ indexes the number of local phonons, $b_q$
($b_q^+$) are the destruction (creation) operators for the phonon $q$
with frequency $\omega_{0q}$. The matrix  $g_{\a\b}$ expresses is the electron-phonon coupling. 
Although feasible, dealing with more than one phonon mode becomes
quickly computationally very demanding, thus in the rest of the this
work we shall consider $N_q=1$. 
In the following we consider a bath discretized into a  number
of bath degrees of freedom and a finite number of available phonons $N_\mathrm{ph}$,
to cut-off the unbounded dimensions of the local phonons Hilbert space.

@@

\subsection{Conserved quantum numbers}\label{sSecQNs}
In order to circumvent the exponential scaling of the dimensions of
the problem scales it is necessary to take into account suitable symmetries.
Indeed, considering operators ${\cal Q}$ such that $[H, {\cal Q}]=0$
reduces of the Fock space into several symmetry sectors
labelled by given quantum numbers $\vec{Q}$. 
In the context of quantum impurity problems two symmetries are often considered: i)
conservation of the total occupation $N$ and ii) the
conservation of the total magnetization $S_z$. Although the total spin
operator $S^2$ can also be conserved, the difficult  implementation
and the marginal gain makes it often non convenient to include this symmetry.
In \NAME we consider three different cases which are controlled by the
input variable {\tt ed\_mode}={\bf normal}, {\bf superc}, {\bf
  nonsu2}.

@@

The {\bf normal} case requires a brief remark. Because of the
independent conservation of $N_\up$ and $N_\dw$, the local Hilbert
space and the electronic Fock can be factorizes as, respectively,
$\HH=\HH_\up\otimes\HH_\dw$, $\FF_\mathrm{e} = \FF_{e\up}\otimes \FF_{e\dw}$.  
Accordingly any Fock state is written as $\ket{\vec{n}_\up}\ket{\vec{n}_\dw}$, the symmetry sector can be written as  $\SS_{\vec{Q}} = \SS_{N_\up}\otimes
\SS_{N_\dw}$ and, correspondingly, the sector map splits in two
mutually exclusive ones $\MM = \MM_\up
\otimes \MM_\dw$.
Thus, each state $\ket{i}=\ket{i_\up}\ket{i_\dw}$ of the
sector is labelled by two integers $[i_\up,i_\dw]$, 
$i_\sigma=1,\dots,D_{\SS_\sigma}$ such that $i=i_\up + i_\dw
D_{\SS_\dw}$. The map $\MM$ connects any such state to a Fock state
$\ket{I}=\ket{I_\up}\ket{I_\dw}$ labelled by two integers
$[I_\up,I_\dw]$ as $I=I_\up +   I_\dw 2^{N_\mathrm{s}}$. For a thorough
discussion about Fock basis organization in this case see
\onlinecite{Amaricci2022}.

The presence of a symmetry induces a factorization of the Fock space,
in turn inducing a block diagonal form to the Hamiltonian matrix.
Each block, labelled by the quantum numbers $\vec{Q}$, has dimension
$D_{\SS(\vec{Q})}$. The sector Hamiltonian matrix $H_\SS$ is represented in the
basis $\ket{i}\in\SS_{\vec{Q}}$ as a sparse matrix. In the {\bf
  normal} case $H_\SS$ takes a particularly symmetric form thanks to
the product structure of the sector and its map~\cite{Amaricci2022}.
The analysis of the spectrum is then reconducted to the inspection of
the Hamiltonian in each symmetry sector. Should particular constraint
holds the search can be limited to only particular sector, further
reducing the computational cost.

@@

%GBtodo placeholder: I passed on this already (except I don't like the minipage of the listings)
\subsubsection{Sparse map}\label{CodeSparseMap}
As it will be evident in the following \secu{sSecRDM}, there are cases in
which the sector construction requires to store, for any sector state,
disjoint information about the corresponding Fock state.
Let's consider a Fock state $\ket{I}$. Its bit decomposition
$\ket{\vec{n}}$ can be further split into arbitrarily grouped chunks of bits.
For instance, we can choose to group the bits into impurity and bath degrees of freedom,
as $\ket{\vec{n}}=\ket{\vec{i}_\up\,\vec{b}_\up\,\vec{i}_\dw\,\vec{b}_\dw}$, where we 
highlight the overall ordering in spin-up and spin-down parts underlying the whole
codebase (see \equ{eq:FockState_|nupndw>} and the subsequent discussion).  
The module {\tt ED\_SPARSE\_MAP} implements an hash-table {\tt sparse\_map}, 
as a component of the {\tt sector} object, which stores for every impurity 
configuration $\vec{i}_\s$ (\emph{key}) all the possible bath states $\vec{b}_\s$ (\emph{values})
that are compatible with the given conserved quantum numbers (see \secu{sSecQNs}).
The {\tt sparse\_map} ${\cal S}$ objects are constructed upon call of the {\tt
  build\_sector} subroutine, automatically dispatched to a different algorithm 
for each value of {\tt ed\_mode}.
In the {\bf normal} case, the sector object contains two {\tt
  sparse\_maps} ${\cal S}_{\s=\up,\dw}$, in line with the
spin dependent factorization of the Fock and sector states: 
$\ket{J}=\ket{J_\up}\otimes\ket{J_\dw} \xleftarrow{\MM}
\ket{j_\up}\otimes\ket{j_\dw}=\ket{j}$.
The sparse maps are built as follows. 
For any spin state of the sector $\ket{J_\s}=\ket{{\vec{i}_{\s}\vec{b}_{\s}}}$,
the \emph{key} is determined by the integer $I_\s$ corresponding to the
impurity bitset $\vec{i}_\s$. The \emph{values} are given by any
integer $B_\s$ corresponding to any bath bitset
$\vec{b_\s}$ associated with $\vec{i}_\s$ by the conservation law of the 
given sector: $\mathtt{sum}(\vec{i}_\s) + \mathtt{sum}(\vec{b}_\s) = \mathtt{popcnt}(I_\s) + \mathtt{popcnt}(B_\s)= N_\s$.
Thus, any given \emph{key-value} combination 
reconstructs an integer $J_\s$ representing a Fock state in the
given spin sector according to the rule $J_s = I_\s +
2^{N_\mathrm{imp}}B_\s$, where $N_\mathrm{imp}$ is the number of impurity bits.

@@

This operations easily takes over more than 80\% of the computation
time, so its optimization is a crucial aspect in any diagonalization
algorithms. In \onlinecite{Amaricci2022} we covered in full details
some distrubuted memory massively parallel algorithms to address this
specific issue.
Here, using the same naming convention introduced in \cite{Amaricci2022}, 
we briefly outline the main strategy for the implementation of the MVP
according to the value of {\tt ed\_mode}.

\paragraph{{\bf normal}}
This case relies on the tensor structure of the Fock space and the
symmetry sectors. In any given sector the electronic part of the
Hamiltonian matrix reads:
\begin{equation}
H_\SS = H_d  + H_\up\otimes \mathbb{I}_\dw + \mathbb{I}_\up\otimes
H_\dw + H_{nd}\,.
\label{HssNormal}
\end{equation}
where $H_d$ is a diagonal term containing the local part of the
Hamiltonian, there including the density-density terms of the
interaction,  $H_\sigma$ describes the hopping processes of the
electrons with spin $\sigma=\up,\dw$.
The term $H_{nd}$ contains all the remaining non-diagonal
elements which do not fit in the previous two components,
e.g. the spin-exchange and pair-hopping interaction terms.
Correspondingly, the any sector state $\ket{v}$ can be expressed in a matrix
basis $\hat{v}$ with rows (columns) running over spin $\up$ ($\down$)
configurations. Leveraging on this setup it is possible the MVP using the 
{\tt MPI\_All2AllV} which aims at optimizing locality in the memory
locality. Additional terms contained in $H_{nd}$ are treated using the
{\tt MPI\_AllGatherV} algorithm which includes a slight communication
overload ultimately reducing the parallel efficiency.

\paragraph{{\bf superc/nonsu2}}
In this two cases the symmetry enforce a specific relation between $\up$ and
$\dw$ configurations ultimately suppressing the benefits arising from
Fock space factorization.
The electronic part of the Hamiltonian matrix is usually organized in this structure:
$$
H_\SS = H_{imp} + H_{int} + H_{bath} + H_{imp-bath} 
$$
where, with almost obvious meaning of the labels, $H_{imp}$ encodes
the non-interacting part of the impurity Hamiltonian determined by
$h^0_{\a\b\s\s'}$, $H_{int}$ the interaction term, $H_{bath}$ the
effective bath terms related to $h^p_{\a\b\s\s'}$ and finally
$H_{imp-bath}$ describes the coupling between impurity and bath
proportional to $V^p_{\a\b\s\s'}$. 
In an MPI setup the first three terms contains elements local in the
memory of each node. These are stored separately from the non-local
ones in any {\tt sparse\_matrix} representing $H_\SS$. 
The MVP is carried out using the {\tt MPI\_AllGatherV} algorithm
which, as discussed in \onlinecite{Amaricci2022}, requires the
reconstruction of the distributed input array any node producing a
overflow in the MPI communication.

@@

The function {\tt vecDim\_Hv\_sector\_?} returns the 
dimension of the vector to be used in the
MVP. In serial execution this number is just $D_\SS$ the dimension of
the sector, whereas in a parallel mode the returned value is the
dimension $d_i$ of the vector chunk per each node such that
$\sum_{i=1}^{N_{nodes}}d_i = D_\SS$. Yet, the specific value of $d_i$
depends on the MPI algorithm used for the MVP function, which in turn
is determined by the symmetry of the problem, i.e. {\tt ed\_mode}.

@@

\subsection{Reduced impurity density matrix}\label{sSecRDM}
This version of the \NAME library features the calculation of the
impurity Reduced Density Matrix (iRDM, $\rho^\mathrm{imp}$), extending 
the algorithm introduced in \onlinecite{BellomiaPhD} so to enable the
analysis of entanglement properties of the quantum impurity for any value
of {\tt ed\_mode}.
To simplify the discussion in this section we take the zero
temperature limit and consider the presence of a non-degenerate
groundstate $\ket{\psi}$ populating the {\tt state\_list}.
The generalization to the presence of degenerate states or the finite
temperature regime are straightforward, as it amounts to defining an
average over degenerate/thermal states $\ibra\rho^\mathrm{imp}\iket$, 
akin to \equ{eq:thermal_average}.
The given pure state belongs to a unique symmetry sector $\SS$ and it
is represented in the Fock basis as $\ket{\psi} = \sum_I a_I \ket{I}$. 
The corresponding pure density matrix $\rho$ is thus explicitly decomposed as:
$$
\rho = \ket{\psi}\bra{\psi} = \sum_{nm=1}^{4^{N_\mathrm{s}}} a^*_ma_n \ket{n}\bra{m} =\sum_{nm=1}^{4^{N_\mathrm{s}}} \rho_{nm} \ket{n}\bra{m}.
$$
The iRDM is obtained from $\rho$ by tracing out the bath degrees of
freedom: $\rho^\mathrm{imp}=\Tr_\mathrm{bath}\rho$. The summations appearing in the
previous expressions become quickly unmanageable with the system
size, as their extrema scale exponentially with $N_\mathrm{s}$. 
To overcome this issue we implemented a fast algorithm
leveraging on the decomposition of the Fock space into symmetry sectors
and the subsequent map sparsity.
Using quantum number conservation the summation extrema get reduced to
the size $D_\SS$ of the sector $\SS$ to which the groundstate (or in
general any eigenstate) belongs. 
Next we analyze the groundstate expansion on the Fock states, splitting the bit representation of any state $\ket{I}$ into
their spin-dependent impurity and bath parts as $\ket{I} = \ket{
  i_\uparrow\,  b_\uparrow\, i_\downarrow \, b_\downarrow}$. Each bitset 
span a space of dimension $D^\a_{imp}$ or $D^\s_{bath}$. 
Finally, as detailed in \secu{CodeSparseMap}, we store into the {\tt sparse\_map} ${\cal S}$, 
for every impurity configuration 
$i_\s =1,\dots,D^\s_{imp}$ (\emph{key}), the bath states 
$j\in[1,\dots,D^\s_{bath}]$ (\emph{values}) that satisfy the conservation law defined
by the \texttt{ed\_mode} variable.
We indicate with ${\cal D}^\s_{i}$ the number of values for the key $i_\s$.
The fast summation algorithm for the evaluation of the iRDM takes on a
different form for the two cases {\bf normal} and {\bf superc/nonsu2},
as outlined below.

>>>>

\subsection{The quantum impurity problem}\label{sSecQIM}
We consider a general quantum impurity problem defined by the
following Hamiltonian:
$$
\hat{H} = \hat{H}_\mathrm{imp} + \hat{H}_\mathrm{bath} + \hat{H}_\mathrm{hyb} + \hat{H}_\mathrm{ph} + \hat{H}_\mathrm{e-ph}
$$
which describes a  multi-orbital interacting quantum
impurity coupled to an electronic bath and to local, i.e. Holstein,
phonons. 
\begin{equation}\label{Himp}
  \begin{split}
    \hat{H}_\mathrm{imp} & = \hat{H}^0_\mathrm{imp} + \hat{H}^{int}_\mathrm{imp}\\
    \hat{H}^0_\mathrm{imp} & =
    \sum_{\a\b\sigma\sigma'}h^{0}_{\a\b\sigma\sigma'}d^{+}_{\a\sigma}d_{\b\sigma'}\\
\end{split}
\end{equation}

@@

\begin{equation}\label{Hint}
  \begin{split}
    \hat{H}^\mathrm{int}_\mathrm{imp} &=U\sum_{\a}n_{\a\uparrow}n_{\a\downarrow}+U'\sum_{\a\neq \b}n_{\a\uparrow}n_{\b\downarrow}+(U'-J)\sum_{\a<\b,\sigma}n_{\a\sigma}n_{\b\sigma}\\
    &{\phantom =}- J_X\sum_{\a\neq
      \b}d^{+}_{\a\uparrow}d_{\a\downarrow}d^{+}_{\b\downarrow}d_{\b\uparrow}+J_P\sum_{\a
      \neq
      \b}d^{+}_{\a\uparrow}d^{+}_{\a\downarrow}d_{\b\downarrow}d_{\b\uparrow}\\
\end{split}
\end{equation}

where $d_{\a\sigma}$ ($d^+_{\a\sigma}$) are the destruction (creation)
second-quantization operators for impurity electrons in the
orbital $\alpha=1,\dots,N_\a$, with $N_\a$ the number of orbitals,
with spin $\sigma=\up,\dw$ and whose occupation is described
by the  operator  
$n_{\alpha\sigma}=d^{+}_{\alpha\sigma}d_{\alpha\sigma}$. 
The non-interacting internal structure of the impurity is described by
the $h^{0}_{\a\b\sigma\sigma'}$ matrix. 
$\hat{H}^\mathrm{int}$ describes the  local multi-orbital
interaction~\cite{Georges2013ACMP} which, for simplicity, we take as a
generalized Hubbard-Kanamori form.  
The first three terms represent the density-density part of the
interaction, where $U$ is the local intra-orbital Coulomb repulsion,
$U'$ the inter-orbital one and $J$ the Hund's coupling~\cite{Georges2013ACMP}.  
The  remaining two terms are, respectively, the spin-exchange and
pair-hopping which we considered with their respective independent
couplings $J_X$ and $J_P$.
In the case $N_\a=3$ a fully symmetric $SU(3)_\mathrm{orbital}\otimes SU(2)_\mathrm{spin}\otimes
U(1)_\mathrm{charge}$ form of the interaction is obtained by setting $U'=U-2J$ and
$J_X=J_P=J$~\cite{Georges2013ACMP}. Different choices, preserving part
of the combined symmetry group, can be made for other values of
$N_\a$~\cite{Georges2013ACMP}.

The bath part and its coupling to the impurity has the form: 
\begin{equation}\label{Hbath}
  \begin{split}
    \hat{H}_\mathrm{bath} &=
    \sum_p\sum_{\a\b\sigma\sigma'}h^p_{\a\b\sigma\sigma'}a^{+}_{p\a\sigma}a_{p\b\sigma'}\\
    %
    \hat{H}_\mathrm{hyb} &= \sum_p\sum_{\a\b\sigma\sigma'}V^p_{\a\b\sigma\sigma'}d^{+}_{\a\sigma}a_{p\b\sigma'}+H.c. \\
\end{split}
\end{equation}
where $p=1,\dots,N_\mathrm{bath}$ is an index running over a finite number of
bath elements, $a_{p\alpha\sigma}$ ($a^+_{p\alpha\sigma}$) are the destruction (creation) operators for
the bath electrons with index $p$, with orbital $\alpha$ and
spin $\sigma$.
The  properties of each bath level are described by the
matrices $h^p_{\a\b\sigma\sigma'}$. As such any bath element can be
composed of several electronic levels according the bath topology,
which will be discussed further in the following.
Each bath level couples to the impurity with an amplitude
$V^p_{\a\b\s\sigma'}$ which we allow to couple different orbital and
opposite spins.

Finally, the electron-phonon part of the quantum impurity problems is
described by the Hamiltonian terms: 
\begin{equation}\label{Hph}
  \begin{split}
    \hat{H}_\mathrm{ph}&=\sum_q \omega_{0q} b_q^+b_q\\
    % 
    \hat{H}_\mathrm{e-ph} &= \sum_q\sum_{\a\b\sigma} g_{\a\b} d^+_{\a\sigma}d_{\b\sigma}(b_q+b_q^+)
\end{split}
\end{equation}
where $q=1,\dots,N_q$ indexes the number of local phonons, $b_q$
($b_q^+$) are the destruction (creation) operators for the phonon $q$
with frequency $\omega_{0q}$. The matrix  $g_{\a\b}$ expresses is the electron-phonon coupling. 
Although feasible, dealing with more than one phonon mode becomes
quickly computationally very demanding, thus in the rest of the this
work we shall consider $N_q=1$. 
In the following we consider a bath discretized into a  number
of bath degrees of freedom and a finite number of available phonons $N_\mathrm{ph}$,
to cut-off the unbounded dimensions of the local phonons Hilbert space.

@@

\subsection{Conserved quantum numbers}\label{sSecQNs}
In order to circumvent the exponential scaling of the dimensions of
the problem scales it is necessary to take into account suitable symmetries.
Indeed, considering operators ${\cal Q}$ such that $[H, {\cal Q}]=0$
reduces of the Fock space into several symmetry sectors
labeled by given quantum numbers $\vec{Q}$. 
In the context of quantum impurity problems two symmetries are often considered: i)
conservation of the total occupation $N$ and ii) the
conservation of the total magnetization $S_z$. Although the total spin
operator $S^2$ can also be conserved, the difficult  implementation
and the marginal gain makes it often non convenient to include this symmetry.
In \NAME we consider three different cases which are controlled by the
input variable {\tt ed\_mode}={\bf normal}, {\bf superc}, {\bf
  nonsu2}.

@@

The {\bf normal} case requires a brief remark. Because of the
independent conservation of $N_\up$ and $N_\dw$, the local Hilbert
space and the electronic Fock can be factorizes as, respectively,
$\HH=\HH_\up\otimes\HH_\dw$, $\FF_\mathrm{e} = \FF_{e\up}\otimes \FF_{e\dw}$.  
Accordingly any Fock state is written as $\ket{\vec{n}_\up}\ket{\vec{n}_\dw}$, the symmetry sector can be written as  $\SS_{\vec{Q}} = \SS_{N_\up}\otimes
\SS_{N_\dw}$ and, correspondingly, the sector map splits in two
mutually exclusive ones $\MM = \MM_\up
\otimes \MM_\dw$.
Thus, each state $\ket{i}=\ket{i_\up}\ket{i_\dw}$ of the
sector is labeled by two integers $[i_\up,i_\dw]$, 
$i_\sigma=1,\dots,D_{\SS_\sigma}$ such that $i=i_\up + i_\dw
D_{\SS_\dw}$. The map $\MM$ connects any such state to a Fock state
$\ket{I}=\ket{I_\up}\ket{I_\dw}$ labeled by two integers
$[I_\up,I_\dw]$ as $I=I_\up +   I_\dw 2^{N_\mathrm{s}}$. For a thorough
discussion about Fock basis organization in this case see
\onlinecite{Amaricci2022}.

The presence of a symmetry induces a factorization of the Fock space,
in turn inducing a block diagonal form to the Hamiltonian matrix.
Each block, labeled by the quantum numbers $\vec{Q}$, has dimension
$D_{\SS(\vec{Q})}$. The sector Hamiltonian matrix $H_\SS$ is represented in the
basis $\ket{i}\in\SS_{\vec{Q}}$ as a sparse matrix. In the {\bf
  normal} case $H_\SS$ takes a particularly symmetric form thanks to
the product structure of the sector and its map~\cite{Amaricci2022}.
The analysis of the spectrum is then reconducted to the inspection of
the Hamiltonian in each symmetry sector. Should particular constraint
holds the search can be limited to only particular sector, further
reducing the computational cost.

@@

%GBtodo placeholder: I passed on this already (except I don't like the minipage of the listings)
\subsubsection{Sparse map}\label{CodeSparseMap}
As it will be evident in the following \secu{sSecRDM}, there are cases in
which the sector construction requires to store, for any sector state,
disjoint information about the corresponding Fock state.
Let's consider a Fock state $\ket{I}$. Its bit decomposition
$\ket{\vec{n}}$ can be further split into arbitrarily grouped chunks of bits.
For instance, we can choose to group the bits into impurity and bath degrees of freedom,
as $\ket{\vec{n}}=\ket{\vec{i}_\up\,\vec{b}_\up\,\vec{i}_\dw\,\vec{b}_\dw}$, where we 
highlight the overall ordering in spin-up and spin-down parts underlying the whole
codebase (see \equ{eq:FockState_|nupndw>} and the subsequent discussion).  
The module {\tt ED\_SPARSE\_MAP} implements an hash-table {\tt sparse\_map}, 
as a component of the {\tt sector} object, which stores for every impurity 
configuration $\vec{i}_\s$ (\emph{key}) all the possible bath states $\vec{b}_\s$ (\emph{values})
that are compatible with the given conserved quantum numbers (see \secu{sSecQNs}).
The {\tt sparse\_map} ${\cal S}$ objects are constructed upon call of the {\tt
  build\_sector} subroutine, automatically dispatched to a different algorithm 
for each value of {\tt ed\_mode}.
In the {\bf normal} case, the sector object contains two {\tt
  sparse\_maps} ${\cal S}_{\s=\up,\dw}$, in line with the
spin dependent factorization of the Fock and sector states: 
$\ket{J}=\ket{J_\up}\otimes\ket{J_\dw} \xleftarrow{\MM}
\ket{j_\up}\otimes\ket{j_\dw}=\ket{j}$.
The sparse maps are built as follows. 
For any spin state of the sector $\ket{J_\s}=\ket{{\vec{i}_{\s}\vec{b}_{\s}}}$,
the \emph{key} is determined by the integer $I_\s$ corresponding to the
impurity bitset $\vec{i}_\s$. The \emph{values} are given by any
integer $B_\s$ corresponding to any bath bitset
$\vec{b_\s}$ associated with $\vec{i}_\s$ by the conservation law of the 
given sector: $\#\vec{i}_\s + \#\vec{b}_\s = \mathtt{popcnt}(I_\s) + \mathtt{popcnt}(B_\s)= N_\s$.
Thus, any given \emph{key-value} combination 
reconstructs an integer $J_\s$ representing a Fock state in the
given spin sector according to the rule $J_s = I_\s +
2^{N_\mathrm{imp}}B_\s$, where $N_\mathrm{imp}$ is the number of impurity bits.

@@

This operations easily takes over more than 80\% of the computation
time, so its optimization is a crucial aspect in any diagonalization
algorithm. In \onlinecite{Amaricci2022} we covered in full detail
the distrubuted memory massively parallel algorithms we implement 
to address this specific issue.
Here, using the same naming convention introduced in \onlinecite{Amaricci2022}, 
we briefly outline the main strategy for the implementation of the MVP
according to the value of {\tt ed\_mode}.

\paragraph{{\bf normal}}
This case relies on the tensor structure of the Fock space and the
symmetry sectors. In any given sector the electronic part of the
Hamiltonian matrix reads:
\begin{equation}
H_\SS = H_\mathrm{d}  + H_\up\otimes \mathbb{I}_\dw + \mathbb{I}_\up\otimes
H_\dw + H_\mathrm{nd}\,.
\label{HssNormal}
\end{equation}
where $H_d$ is a diagonal term containing the local part of the
Hamiltonian, there including the density-density terms of the
interaction,  $H_\sigma$ describes the hopping processes of the
electrons with spin $\sigma=\up,\dw$.
The term $H_{nd}$ contains all the remaining non-diagonal
elements which do not fit in the previous two components,
e.g. the spin-exchange and pair-hopping interaction terms.
Correspondingly, the any sector state $\ket{v}$ can be expressed in a matrix
basis $\hat{v}$ with rows (columns) running over spin $\up$ ($\down$)
configurations. Leveraging on this setup it is possible the MVP using the 
{\tt MPI\_All2AllV} which aims at optimizing locality in the memory
locality. Additional terms contained in $H_{nd}$ are treated using the
{\tt MPI\_AllGatherV} algorithm which includes a slight communication
overload ultimately reducing the parallel efficiency.

\paragraph{{\bf superc/nonsu2}}
In this two cases the symmetry enforce a specific relation between $\up$ and
$\dw$ configurations ultimately suppressing the benefits arising from
Fock space factorization.
The electronic part of the Hamiltonian matrix is usually organized in this structure:
$$
H_\SS = H_\mathrm{imp} + H_\mathrm{int} + H_\mathrm{bath} + H_\mathrm{imp-bath} 
$$
where, with almost obvious meaning of the labels, $H_\mathrm{imp}$ encodes
the non-interacting part of the impurity Hamiltonian determined by
$h^0_{\a\b\s\s'}$, $H_\mathrm{int}$ the interaction term, $H_\mathrm{bath}$ the
effective bath terms related to $h^p_{\a\b\s\s'}$ and finally
$H_\mathrm{imp-bath}$ describes the coupling between impurity and bath
proportional to $V^p_{\a\b\s\s'}$. 
In an MPI setup the first three terms contains elements local in the
memory of each node. These are stored separately from the non-local
ones in any {\tt sparse\_matrix} representing $H_\SS$. 
The MVP is carried out using the {\tt MPI\_AllGatherV} algorithm
which, as discussed in \onlinecite{Amaricci2022}, requires the
reconstruction of the distributed input array any node producing a
overflow in the MPI communication.

@@

The function {\tt vecDim\_Hv\_sector\_?} returns the 
dimension of the vector to be used in the
MVP. In serial execution this number is just $D_\SS$ the dimension of
the sector, whereas in a parallel mode the returned value is the
dimension $d_i$ of the vector chunk per each node such that
$\sum_{i=1}^{N_\mathrm{nodes}}d_i = D_\SS$. Yet, the specific value of $d_i$
depends on the MPI algorithm used for the MVP function, which in turn
is determined by the symmetry of the problem, i.e. {\tt ed\_mode}.

@@

\subsection{Reduced impurity density matrix}\label{sSecRDM}
The new \NAME version of the library features the calculation of the
impurity Reduced Density Matrix (iRDM, $\rho^\mathrm{imp}$), extending 
the algorithm introduced in \onlinecite{BellomiaPhD} so to enable the
analysis of entanglement properties of the quantum impurity for any value
of {\tt ed\_mode}.
To simplify the discussion in this section we take the zero
temperature limit and consider the presence of a non-degenerate
groundstate $\ket{\psi}$ populating the {\tt state\_list}.
The generalization to the presence of degenerate states or the finite
temperature regime are straightforward, as it amounts to defining an
average over degenerate/thermal states $\ibra\rho^\mathrm{imp}\iket$, 
akin to \equ{eq:thermal_average}.
The given pure state belongs to a unique symmetry sector $\SS$ and it
is represented in the Fock basis as $\ket{\psi} = \sum_I a_I \ket{I}$. 
The corresponding pure density matrix $\rho$ is thus explicitly decomposed as:
$$
\rho = \ket{\psi}\bra{\psi} = \sum_{IJ=1}^{4^{N_\mathrm{s}}} a^*_J a_I \ket{I}\bra{J} =\sum_{IJ=1}^{4^{N_\mathrm{s}}} \rho_{IJ} \ket{I}\bra{J}.
$$
The iRDM is obtained from $\rho$ by tracing out the bath degrees of
freedom: $\rho^\mathrm{imp}=\Tr_\mathrm{bath}\rho$. The summations appearing in the
previous expressions become quickly unmanageable with the system
size, as their extrema scale exponentially with $N_\mathrm{s}$. 
To overcome this issue we implemented a fast algorithm
leveraging on the decomposition of the Fock space into symmetry sectors
and the subsequent map sparsity.
Using quantum number conservation the summation extrema get reduced to
the size $D_\SS$ of the sector $\SS$ to which the groundstate (or in
general any eigenstate) belongs. 
Next we analyze the groundstate expansion on the Fock states, splitting the bit representation of any state $\ket{I}$ into
their spin-dependent impurity and bath parts as $\ket{I} = \ket{
  i_\uparrow\,  b_\uparrow\,} \otimes \ket{i_\downarrow \, b_\downarrow}$. Each bitset 
span a space of dimension $D^\a_\mathrm{imp}$ or $D^\s_\mathrm{bath}$. 
Finally, as detailed in \secu{CodeSparseMap}, we store into the {\tt sparse\_map} ${\cal S}$, 
for every impurity configuration 
$i_\s =1,\dots,D^\s_{imp}$ (\emph{key}), the bath states 
$b_\s\in[1,\dots,D^\s_\mathrm{bath}]$ (\emph{values}) that satisfy the conservation law defined
by the \texttt{ed\_mode} variable.
We indicate with ${\cal D}_{i\s}$ the number of keys $i_\s$ and with 
${\cal D}_{b\s}$ the number of values $b_\s$.
The fast summation algorithm for the evaluation of the iRDM takes on a
different form for the two cases {\bf normal} and {\bf superc/nonsu2},
as outlined below.

\paragraph{{\bf normal}}
In this case we take advantage of the Fock space and sector spin-resolved
factorization, discussed in depth in \secu{sSecQNs}. We can indeed split the
global bitset in spin-resolved impurity and bath states, as
